---
title: "BroadbandNow Open Data Exploration"
author: "Cary"
date: "2/6/2021"
output:
  github_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Import R packages/libraries

```{r import libraries}
library(tidyverse)
library(ggplot2)
library(readr)
library(readxl)
library(stringr)
library(janitor)
```

## Load and seperate the dataset into 2019 and 2015 providers

```{r}
#Retrieve broadband data from Github
broadbandurl <- 'https://raw.githubusercontent.com/BroadbandNow/Open-Data/master/broadband_data_opendatachallenge.csv'
broadband <-read.csv(url(broadbandurl), header = T, sep = ",", strip.white = TRUE)
# The raw data have numerical values in population, percent of access and lowest price.
# However, they were loaded as character and needs to be converted to numeric 
```

Even though the column names indicated 2020, the actual data is from June 2019 FCC 477 reports. The zipcode, county, state information are kept in both files for reference. 
```{r}
head(broadband) #Load dataframe

colnames(broadband) #Check column names and index
get_dupes(broadband) #Check for duplicates with janitor package
```
## Rename selected column and convert data type 

```{r}
colnames(broadband)[16] <- 'Month_price' #Rename the lowest monthly price for internet only service in wired and fixed wireless
numeric(length = 2) #Set the option to avoid coercing values into integer
broadband$Month_price <- as.numeric(broadband$Month_price) #Convert values to float

colnames(broadband)[15]<- 'Percent_access' #Rename the percentage of access to fixed and fixed wireless broadband
#Parse all the "%" character from the Percent-access column
broadband$Percent_access <- str_replace(broadband$Percent_access, "\\%", "")

broadband$Percent_access <- as.numeric(broadband$Percent_access) #Convert values to numeric
```

```{r}
broadband$Population <- as.numeric(broadband$Population) # Convert Population to numeric

broadband$Zip <- as.character(broadband$Zip) #Convert Zipcode to character 

#The All100_3 columns are missing the proper year 
colnames(broadband)[11] <- 'All100_3_2020' #Rename the 11th column to add 2020
colnames(broadband)[23] <- 'All100_3_2015' #Rename the last column to add 2015
```


## Descriptive and Missing Values 

```{r}
summary(broadband)
```

## Using Dallas County as an example, I'm assessing how does the does the data looks like for a metropolitan area 
```{r}
dallas <- broadband %>% filter(County == 'Dallas')
# this return all rows with Dallas county regardless of which state
dallas
```
```{r}
dallas <-broadband %>% filter(State == 'Texas') %>% filter (County == 'Dallas')
#This filter by state and then by county to get only Dallas County data within Texas

dallas
```
```{r}
# As you noticed, there are quiet a few of missing information per zip code within Dallas County in Texas
summary(dallas)

```
38 rows missing within 122 rows is about 31%

After checking the postal codes in [City-Data.com](https://www.city-data.com/zipDir.html), for example 75391-75399, this set of zip code doesn't have any population information associated with it. 

## Using a rural area to check the dataset. The census data has a [Rural America](https://mtgis-portal.geo.census.gov/arcgis/apps/MapSeries/index.html?appid=49cd4bc9c8eb444ab51218c1d5001ef6#:~:text=The%20Census%20Bureau%20defines%20rural,tied%20to%20the%20urban%20definition) page for county rurality. 

```{r}
# Crockett County, Texas with a 22.7 Percent Rural 
crockett <- broadband %>% filter(County == 'Crockett')
crockett
#There are five rows filter and with only one row for Texas 
```

```{r}
# Schleicher County, Texas with a 100 Percent Rural 
schleicher <- broadband %>% filter(County == 'Schleicher')
schleicher
#There is only one result for this Schleicher in the dataset 
```

## Slice the data by the years 2015 and 2019

```{r}
#Extract the 2019 broadband information from the pricing and speed information 
data2019 <- broadband[, c(1:16)]
data2015 <- broadband[, c(1:4, 17:23)]

```

## We would like to add a new column to the 2019 dataset considering the true price that a user will pay for internet services including taxes and fees. Currently we estimated a 25% on top of the internet charge as taxes and fees

```{r}
# Append a column to the dataset with consideration of fees and tax on final price 
data2019$est_final_price <- broadband$Month_price * 0.25
```

## Reivew information in the extracted 2019 dataset 

```{r}
colnames(data2019)
```
```{r}
summary(data2019)
```


# Census B28005 Table reformmated version import and conduct futher cleaning 

```{r}
b28005 <- readxl::read_excel('./Datasets/b28005-5yr estimates-geo.xlsx')

head(b28005)
```


```{r}
#Check for missing values - incomplete cases 
b28005[!complete.cases(b28005),]
```

```{r}
#remove those incomplete cases, from 829 row to 826 
b28005 <- na.omit(b28005)
```


```{r}
#Convert County and State as Factors first 
b28005$State <- as.factor(b28005$State)
names(b28005)[1] <- "County" #rename the column to remove random characters
b28005$County <- as.factor(b28005$County)
#Checking the levels in the State variable 
levels(b28005$State)
```


```{r}
#Review data description 
summary(b28005)
```



```{r}
# Need to convert some of the values into numeric form for further calculation
b28005$Under.18.years <- as.integer(b28005$Under.18.years)

headers <- c("Total", "Under.18.years", "Has.a.computer", "With.dial.up.alone", "With.a.broadband","Without.any.internet","No.computer")
b28005[,headers] <- lapply(b28005[,headers], as.numeric, na.rm = F)
```


## Add columns of percentage calculation for under 18 that owns a computer , had broadband

```{r}
b28005$k12pop <- b28005$Under.18.years/b28005$Total
b28005$own_comp_percent <- b28005$Has.a.computer/b28005$Under.18.years
b28005$has_broadband_percent <- b28005$With.a.broadband/b28005$Under.18.years
```

```{r}
b28005
```
## Groupby states for a state dataframe 

```{r}
#using groupby statement, I can go ahead and aggregate the columns that I wanted. 
state_b28005 <- b28005 %>% 
  group_by(State) %>% 
  summarize(state_total = sum(Total), state_k12 = sum(Under.18.years), broadband = sum(With.a.broadband), computer = sum(Has.a.computer))

state_b28005
```

```{r}
state_b28005$k12_percent <- state_b28005$state_k12/state_b28005$state_total
state_b28005$computer_percent <- state_b28005$computer/state_b28005$state_k12
state_b28005$broadband_percent <- state_b28005$broadband/state_b28005$state_k12
```

```{r}
state_b28005
```

