---
title: "BroadbandNow Data Processing"
author: "Cary"
date: "2/6/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Import R packages/libraries

```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(readr)
library(stringr)
library(janitor)
```

# Load dataset with 2019 and 2015 provider information

```{r}
#Retrieve broadband data from previous extraction
broadbandnow <-read.csv("./Datasets/broadbandnow-export.csv", header = T, sep = ",", strip.white = TRUE)
```
View dataframe
```{r}
head(broadbandnow)
```
Column Names
```{r}
colnames(broadbandnow)
```
View data type of each column
```{r}
str(broadbandnow)
```
Even though the column names [5:11] indicated 2020, the actual data is from June 2019 FCC Form 477 reports. The zip-code, county, state information are kept in both files for reference. 
```{r eval=FALSE, include=FALSE}
get_dupes(broadbandnow) # No duplicates, check with janitor package
```
## Rename column names and convert data type 

```{r}
colnames(broadbandnow)[16] <- 'Month_price' #Rename the lowest monthly price for internet only service in wired and fixed wireless

numeric(length = 2) #Set the option to avoid coercing values into integer

broadbandnow$Month_price <- as.numeric(broadbandnow$Month_price) #Convert values to float

colnames(broadbandnow)[15] <- 'Percent_access' #Rename the percentage of access to fixed and fixed wireless broadband 

#Parse all the "%" character from the Percent-access column
broadbandnow$Percent_access <- str_replace(broadbandnow$Percent_access, "\\%", "")

broadbandnow$Percent_access <- as.numeric(broadbandnow$Percent_access) #Convert values to numeric
```
View data type again
```{r}
str(broadbandnow)
```


```{r}
broadbandnow$Population <- as.numeric(broadbandnow$Population) # Convert Population to numeric

broadbandnow$Zip <- as.character(broadbandnow$Zip) #Convert Zipcode to character 

#The All100_3 columns are missing the proper year 
colnames(broadbandnow)[11] <- 'All100_3_2020' #Rename the 11th column to add 2020
colnames(broadbandnow)[23] <- 'All100_3_2015' #Rename the last column to add 2015
```
View data type again 
```{r}
str(broadbandnow)
```
Note. %Access to Terrestrial Broadband (renamed to Percent_access): Percent of the Zip's Population that has Access to Terrestrial (Wired + Fixed Wireless) Broadband (25 Mbps Download / 3 Mbps Upload)

## Descriptive and Missing Values 
There are 2881 zip-code without any population information 
There are 2923 zip-code without Percent access information
```{r}
summary(broadbandnow)
```
Determine which zip-code have missing data 
```{r}
# There are 32608 rows of zip-code loaded in the dataframe 
# The number of incomplete cases are 8389 (32608-24219) 
sum(complete.cases(broadbandnow))
```
# Split the two years 
The missing values could be because of the two years in the original dataframe 

```{r}
# Both subset will contain zip-code, county information, and population 
broadbandnow_2019 <- broadbandnow[, c(1:16)]
broadbandnow_2015 <- broadbandnow[, c(1:4, 17:23)]
```
Note. Population: Total Population of Zip Code (2010) in Broadband-Now 
## 2015
```{r}
sum(complete.cases(broadbandnow_2015))
```
There are 29724 complete cases in 2015.

The 2015 data only has information about the providers. There is no percent access, or speed
information with it. 
## 2019
```{r}
sum(complete.cases(broadbandnow_2019))
```
There are 24219 complete cases in 2019, which is 5505 difference to 2015
### Zip-code with missing information 
Determine which zip-code have missing values, by producing a dataframe 
```{r}
broadbandnow_2019[!complete.cases(broadbandnow_2019),] # 8389 rows
```


```{r}
popu_2019 <- read.csv("./Datasets/ACS_5Y2019.B01003_popu.csv", header = T)
str(popu_2019)
```
Split the Geographic.Area.Name into two columns first because there are two codes in it 
```{r}
popu_2019 <- separate(popu_2019, col = Geographic.Area.Name, into = c("ZCTAs", "Zip"), sep = " ")
```

Join the population data to the Broadband-now dataframe 
```{r}
data <- broadbandnow_2019 %>% left_join(popu_2019, by = "Zip")
```
The output has 32608 rows and the added Census information can verify if the zipcode
has any people living there, or simply a postal code. 

```{r}
data <- data %>% drop_na(c("Population", "Estimated_total"))
```
There are 27366 rows of zip-code left 

```{r}
sum(complete.cases(data))
```

```{r eval=FALSE, include=FALSE}
# Export this level of data 
write.csv(data, "./Datasets/broadbandnow-2019.csv", row.names = F)
```




