---
title: "Aggregate Locale and Title I Status by State"
author: Cary K. Jim
output: html_notebook
---
```{r eval=FALSE, include=FALSE}
# This has been published to the RPubs as NCES Public School 2018-2019 Locale and Title I status file on Oct 26,2021
```


Import Library/Packages

```{r}
library(tidyverse)
```

Load the processed public school file from Part 2 <https://rpubs.com/caryjim/nces-ps-data-processing>

```{r Load dataset}
public_school <- read.csv("./Datasets/public_school_processed.csv", header = T, stringsAsFactors = T)
# This time, stringsAsFactors is applied to ensure all characters are converted
```

View columns

```{r column names}
colnames(public_school)
```

View data type

```{r data type}
str(public_school)
```

## State Level

In the original public school dataset, state names entries are inconsistent. There are some state names as ALL CAPS and some as lower cases. Therefore, using state abbreviation is a better varaible unless we parse the chartacters in the full state names.

```{r Number of school by ST}
# A quick count of schools by each state in the processed data
table(public_school$ST)
```

Order the number of school per state

```{r sort the table by ST}
sort(table(public_school$ST))
```

Note. It would be interesting to see the number of public school per state and compare it to the current digital opportunity status of these regions.

### Locale Information
```{r Number of School by Urbanicity/Rurality}
table(public_school$Urban.centric.Locale)
```

```{r Number of School by Titile I Status}
table(public_school$Title.I.School.Status)
```
Aggregate the types of Locale by State and Provide a Count 
```{r Urbanicity/Rurality by State}
# Collapse the data by State and their locale 
public_school %>% 
  group_by(ST, Urban.centric.Locale) %>%
  summarize(locale = n())
```

Review the Title 1 information by state

```{r Title I School Status by State}
public_school %>% 
  group_by(ST, Title.I.School.Status) %>%
  summarize(each_count = n())
```
We can see that some states have more reported categories of Title I status 
```{r Total enrollment of K-12 student by state}
public_school %>% 
  group_by(ST) %>%
  summarize(total_enrollment = sum(Members))
```

## Task 1. Create a new dataframe for state level information

Group selected variables by state level and calculate a percentage of that information or a total count

### Total student enrollment and gender

```{r Aggregate enrollment info by State}
# Create a new dataframe with state level information 
schools_state <-
  public_school %>% 
  group_by(ST) %>%
  summarise(Num_of_school = n(), Total_enrollment = sum(Members), 
            male = sum(Male, na.rm = TRUE), Female = sum(Female, na.rm = TRUE))
```

Beware that some schools have missing values for one of the gender and the use of sum() of those return a NA value. So we have to use a different method by adding na.rm = True.

```{r view dataframe}
head(schools_state) #There are 51 rows
```

### Locale

```{r Locale count by State}
public_school %>% 
  group_by(ST) %>%
  summarise(Urbanicity = n_distinct(Urban.centric.Locale))
# This method gives me the count of categories, but not how many within the category 
```

```{r Create a dataframe for locale info}
t1 <- public_school %>% 
  group_by(ST, Urban.centric.Locale) %>%
  summarise(Locale_count = n())
t1
```

```{r Reshape dataframe to spread levels as column headings}
t1<- spread(t1, Urban.centric.Locale, Locale_count)
t1
```

```{r Fill-in missing values with 0}
#Row number 8 and row 40 in the t1 dataframe 
t1[8, 4] <- 0	
t1[8, 5] <- 0
t1[40, 5] <- 0

t1
```

```{r Total acorss locale levels}
#Create a total column
t1$total <- t1$City + t1$Suburb + t1$Town + t1$Rural
#Replace each locale by its own percentage 
#Create a list first to check if calculation are done correctly
t1$City <- t1$City/t1$total
```

```{r Replace each level values as percentages}
#Replace each locale by its own percentage 
t1$Suburb <- t1$Suburb/t1$total
t1$Town <- t1$Town/t1$total
t1$Rural <- t1$Rural/t1$total
t1
```
Note. The percentage of each locale by school types can also be used as a comparison to digital opportunity. 

## Task 2 Title I status

```{r Create a dataframe for Title I status levels}
t2 <- public_school %>% 
  group_by(ST, Title.I.School.Status) %>%
  summarise(count = n())
```

```{r}
head(t2)
```


```{r}
levels(t2$Title.I.School.Status)
```


```{r rename factors level}
levels(t2$Title.I.School.Status) <- c("TAE_No", "TAE", "SW_TAE", "SW_No", "SW", "NOT_TitleI", "Unknown")
```

```{r}
head(t2)
```

```{r}
t2 <- spread(t2, Title.I.School.Status, count)
t2
```
Since we have to perform calculation, we will impute all NA as zero 
```{r Fill-in missing counts as 0}
t2[is.na(t2)] = 0
```

```{r Create total column}
t2$total <- rowSums(t2[,2:8])
t2
```


```{r Calculate percentage of each levels by total}
t2$Unknown <- t2$Unknown/t2$total
t2$TAE_No <- t2$TAE_No/t2$total
t2$TAE <- t2$TAE/t2$total
t2$SW_TAE <- t2$SW_TAE/t2$total
t2$SW_No <-t2$SW_No/t2$total
t2$SW <- t2$SW/t2$total
t2$NOT_TitleI <- t2$NOT_TitleI/t2$total
```

Verify the calculation

```{r Verify the calculation}
#Confirming the percentages are calculated correctly
public_school %>% 
  group_by(ST, Title.I.School.Status) %>%
  summarise(each_count = n()) %>%
  mutate(Title1_percent = each_count/sum(each_count))
```

## Task 3. Combine dataframes 

Merge schools_state with urbanciity/rurality table

```{r}
schools_state <- left_join(schools_state, t1, by = "ST")
```

```{r}
head(schools_state)
```

```{r}
#We can drop the total count for urbanicity/rurality -column index 10
schools_state <- schools_state[-c(10)]

head(schools_state)
```

Merge schools_state with Title 1 info

```{r}
#Append the title 1 information by each state 
schools_state <- left_join(schools_state, t2, by = "ST")
```

```{r}
head(schools_state)
```

```{r drop total column (duplicated count)}
#Again, drop the total count column , index 17
schools_state <- schools_state[-c(17)]
```

```{r}
head(schools_state)
```

```{r eval=FALSE, include=FALSE}
#Export the dataframe for record
#write.csv(schools_state,"D:/Documents/R/Digital Divide/Open_Data_Challenge/Analysis/schools_by_state.csv")
```

